### 2024-10-03 Chapters/Chapter4.tex
Updated tables in 5 and added 6

\chapter{Model architecture} % Main chapter title

\label{Chapter4} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}

%----------------------------------------------------------------------------------------
%	SECTION 1
%----------------------------------------------------------------------------------------

@@

\begin{table}[h]
    \centering
    \caption[Hyperparameter search with best performing combination.]{Hyperparamter search with best performing combination shown. Hyperparameter search was done using hyperband algorithm that initially searches randomly for the best parameters but then hones in on what is working and as such is neither exhaustive nor completely random. This means that an upper limit will not be set on the number of combinations to try like with randomsearch.}
    \label{table:gridSearchHyperparamters}
    \begin{tabular}{ccc}
        Parameter & Range of values & Selected\\
        Layers &  min\_value = 4, max\_value = 15, step = 1 & \textbf{10}\\
        Units &  min\_value = 32, max\_value = 512, step = 32 & \textbf{64}\\
        Penalties & min\_value = 1e-5, max\_value = 1, sampling = log & \textbf{1e-4}\\
        Epochs & min\_value = 10, max\_value = 1000, step = 10 & \textbf{50}\\
        Optimizers & Adam, RMSprop, Adamax & \textbf{Adamax}\\
        Activation & ReLU, ELu, Softmax & \textbf{ReLU}\\
    \end{tabular}
\end{table}

>>>>

\chapter{Model Architecture and Training}

\label{Chapter4}

@@

\begin{table}[h]
    \centering
    \caption[Hyperparameter search with best performing combination.]{Hyperparamter search with best performing combination shown. Hyperparameter search was done using hyperband algorithm that initially searches randomly for the best parameters but then hones in on what is working and as such is neither exhaustive nor completely random. This means that a hard upper limit will not be set on the number of combinations to try like with randomsearch.}
    \label{table:gridSearchHyperparamters}
    \begin{tabular}{ccc}
        \toprule
        Parameter & Range of values & Selected\\
        \midrule
        Layers &  min\_value = 4, max\_value = 15, step = 1 & \textbf{10}\\
        Units &  min\_value = 32, max\_value = 512, step = 32 & \textbf{64}\\
        Penalties & min\_value = 1e-5, max\_value = 1, sampling = log & \textbf{1e-4}\\
        Epochs & min\_value = 10, max\_value = 1000, step = 10 & \textbf{250}\\
        Optimizers & Adam, RMSprop, Adamax & \textbf{Adamax}\\
        Activation & ReLU, ELu, Softmax & \textbf{ReLU}\\
        \bottomrule
    \end{tabular}
\end{table}

As mentioned, hyperband doesn't set a hard upper limit to the number of epochs it will train in total. When using the \href{https://keras.io/api/keras_tuner/tuners/hyperband/}{hyperband class} several factors can be set. One of interest here is the $hyperband_iterations$ argument. This determines how often the hyperband algorithm is run and defaults to 1. For each iteration the epochs are distributed between tries (that is each set of hyperparameters) with the total amount of epochs approximately $n_{epochs} = max_{epochs} * log^2(max_{epochs})\approx 10^4$, where $max_{epochs} = 1000$ gives the maximum number of epochs that one set of hyperparameters can be trained for. Searching a space generally takes a lot of time but this drastically improves on gridsearch. If each epoch takes around 10 seconds to run then the total search would take around 28 hours on a shared resource. This is resource intensive and cannot be repeated often. Another question that remains is whether the ranges given are optimal. Are a 1000 epochs even enough?

\section{Model Training}
To determine the required number of epochs a high number of epochs can be selected and the loss plotted. These plots can be seen in Figures (\ref{fig:loss_plot_wo_elevation}, \ref{fig:loss_plot_w_elevation}). Looking at these two plots, training loss decreases for over 1900 epochs in both plots, while the validation loss stops decreasing at around 200 epochs for DEM model and 1800 epochs for model without DEM. This is unexpected. The DEM model has vastly more parameters and as such should contain more useful information to be learned, which one might expect to take longer to learn. This does not seem to be the case. The DEM model only gets trivially better after 1000 epochs while the non DEM model keeps meaningfully improving until around 1800 epochs. With respect to time constraints of shared resources models were trained for 250 epochs.

\begin{figure}
    \centering
    \includegraphics[scale = 0.6]{Figures/loss_plot_wo_elevation.png}
    \caption[Loss plot of model trained for 2000 without DEM.]{Loss plot of model trained for 2000 without DEM.}
    \label{fig:loss_plot_wo_elevation}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[scale = 0.6]{Figures/loss_plot_w_elevation.png}
    \caption[Loss plot of model trained for 2000 with DEM.]{Loss plot of model trained for 2000 with DEM.}
    \label{fig:loss_plot_w_elevation}
\end{figure}

### 2024-10-03 Chapters/Chapter5.tex
Updated tables in 5 and added 6

\begin{table}[h]
    \caption[Model results for different AWSL]{MAPE for each average wind speed limit with and without landscape elevation in a 30°sector around the point of interest into the direction of the reanalysis wind. The influence of adding elevation data seems to reduce the error. The percentage error is higher for lower wind speeds and thus observing the error for different lower bound of wind speed will produce different results. How to determine this lower bound can done by measurements (f) or reanalysis wind speed at 15 m ($ws_{15}$). Using the measured wind will lead to data leakage and thus gives better results. This should not be used generally but is included here to show the difference in performance.}
    \label{table:results}
    \centering
    \begin{tabular}{lcccccc}
        \textbf{AWSL} & \textbf{MAPE} & & & & &\\ 
         & \multicolumn{2}{c}{\textit{Baseline}} &  \multicolumn{2}{c}{\textit{Without DEM}} & \multicolumn{2}{c}{\textit{With DEM}} \\\hline
        m/s & $ws_{15}$ & f & $ws_{15}$ & f & $ws_{15}$ & f\\\hline
        $\geq 0$ & 39.2\% & 39.2\% & 19.7\% & 15.9\% & 18.6\% & 15.3\%\\
        $\geq 5$ & 28.1\% & 14.8\% & 15.9\% & 9.9\% & 14.1\% & 9.4\%\\
        $\geq 10$ & 23.9\% & 11.1\%& 13.6\% & 7.6\% & 11.5\% & 7.4\%\\
        $\geq 15$ & 23.2\% & 9.3\% & 13.0\% & 7.0\% & 10.6\% & 6.4\%\\
        $\geq 20$ & 24.7\% & 8.2\% & 13.6\% & 6.3\% & 11.1\% & 5.8\%\\
        $\geq 25$ & 27.7\% & 7.3\% & 15.7\% & 5.8\% & 12.9\% & 5.5\% \\
    \end{tabular}
\end{table}

%7.5 & 25.3\% & 12.6\% & 14.5\% & 8.6\% & 12.5\% & 8.6\%\\
%12.5 & 23.3\% & 10.1\% & 13.1\% & 7.1\% & 11.0\% & 6.9\%\\
%17.5 & 23.7\% & 8.7\% & 12.8\% & 6.4\% & 10.5\% & 6.1\%\\
%22.5 & 26.1\% & 7.8\% & 14.5\% & 6.1\% & 12.6\% & 5.7\%\\

This is some improvement upon the baseline error, with a decrease in error from 23.9\% to 1.8\% depending on the setup used. The power generated by wind mill increases with wind speed cubed\cite{wind_power}. The highest wind gusts in Iceland are around 70 m/s. Knowing the gust factor to 2 percentage points better than before can allow for the anticipated power generated by wind mills to increase significantly.

Another way to look at the error improvement is by station. No location data was directly included in the training data. In Table \ref{table:station_mae_distribution} the mean absolute error of predicted average wind speed and measured average wind speed can be seen for the extreme values. A question to ask is it possible to achieve better results when only looking at a single station?

\begin{table}[h]
    \caption[Model result by station]{The MAPE results for selected stations of interest, both when training for the specific site and when the stations are a part of the general data. For every station the AWSL is set at 10 m/s. In training for a single station at a time, some site specific information can be gauged. This does not mean that the a better result can be reached for that site. Factors such as the number of datapoints at given location can significantly impact the result. This table uses the measured wind speed to determine the cut off for data points. This leads to some data leakage and an increased performance compared to using the reanalysis CARRA speed for cut off.}
    \label{table:specific_sites}
    \centering
    \begin{tabular}{l  ccc}
        \textbf{Station name} & \textbf{Number of measurement points} & \multicolumn{2}{c}{\textbf{MAPE}}\\
        & & General training & Site training\\\hline
        Akrafjall & 42,791 & 18.6\% & 93.7\%\\
        Almannaskarð & 4,014 & 12.2\% & 86.7\%\\
        Ásgarðsfjall & 15,121 & 9.1\% & 9.4\%\\
        Jökulheimar & 17,176 & 7.7\% & 7.7\%\\
        Sandbúðir & 18,718 & 6.8\% & 6.4\%\\
        Stórholt & 35,126 & 7.1\% & 29.2\% \\
        Þúfuver & 19,538 & 6.4\% & 6.8\%\\
    \end{tabular}
\end{table}

@@

\begin{figure}[h]
    \centering
    \caption[MAPE error distribution of stations shown on a map of Iceland.]{The MAPE error of each station in data shown as color gradient circles. That is each station is represented by one circle, with the error value represented as a color gradient from dark blue to yellow. The lowest error at a single station was around 7\% at Garðskagaviti and the highest around 35\% at Seljalandsdalur. It is important to note that the model is trained using a cutoff of 10 m/s and this cutoff point is determined using the reanalysis wind at 15 meters above ground.}
    \includegraphics[scale = 0.5]{Figures/errorMap.png}
    \label{fig:errorMap}
\end{figure}

@@

Table (\ref{table:closed_intervals}) shows no improvement over the values in the last column in Table (\ref{table:results}). As previously mentioned, the gust factor decreases with increasing wind speed and thus, training on intervals and lessening this effect might be expected to give better results. This does not seem to be the case.

Some interesting sites to look closer at for drivers might include places like Kjalarnes, Ingólfsfjall, Þrengsli and others. These can be seen in Table (\ref{table:specific_sites}).

\begin{table}[h]
    \caption[Model result by stations of interest]{The MAPE results of different stations for several stations of interest, both when training for the specific site and when the stations are a part of the general data. For every station the AWSL is set at 10 m/s. In training for a single station at a time, some site specific information can be gauged. This does not mean that the a better result can be reached for that site. Factors such as the number of datapoints at given location can significantly impact the result. This table uses the measured wind speed to determine the cut off for data points. This leads to some data leakage and an increased performance compared to using the reanalysis CARRA speed for cut off.}
    \label{table:specific_sites}
    \centering
    \begin{tabular}{l | cc}
        \textbf{Station name} & \multicolumn{2}{c}{\textbf{MAPE}}\\
         & Baseline & Model\\\hline
        Fáskrúðsfjörður & 28.2\% & 21.8\%\\
        Ingólfsfjall & 30.0\% & 19.6\%\\
        Kjalarnes & 20.7\% & 13.5\% \\
        Sandskeið & 13.0\% & 10.2\%\\
        Seyðisfjörður & 32.1\% & 23.0\%\\
        Þjórsárdalur & 12.2\% & 11.4\%\\
        Þrengsli & 13.6\% & 11.3\%\\
    \end{tabular}
\end{table}

\begin{figure}[h]
    \centering
    \caption[Feature importance for a single observation of a neural network.]{Feature importance of a neural network with model architecture as described in Table \ref{table:gridSearchHyperparamters} and data as described in Table \ref{table:trainDataExample}. In this specific instance the wind direction ($wd_{15}$) has the highest negative influence and the temperature has the highest positive influence. Elevation data is excluded when working with Shapley values, as the contribution of each elevation point is very low and there are very many of them. To see their influence on the model output see Table \ref{table:results}.}
    \includegraphics[scale = 0.6]{Figures/shap_plots/waterfall_plot.png}
    \label{fig:ShapleyWaterfall}
\end{figure}

In Figure (\ref{fig:ShapleySummary}) the contribution of each feature, excluding the elevation points, can be seen for the model in general (a significant subset of data is used).

\begin{figure}
    \centering
    \caption[Summary feature importance of a neural network.]{Feature importance of a neural network with model architecture as described in Table \ref{table:gridSearchHyperparamters} and data as described in Table \ref{table:trainDataExample}. We can see that generally multiple factors influence the prediction, with the station elevation being highly influential. There is seemingly one outlier for the Richardson number, which usually has very little influence. Elevation data is excluded when working with Shapley values, as the contribution of each elevation point is very low and there are very many of them. To see their influence on the model output see Table \ref{table:results}.}
    \includegraphics[scale = 0.6]{Figures/shap_plots/summary_plot.png}
    \label{fig:ShapleySummary}
\end{figure}

Looking at Figure (\ref{fig:ShapleySummary}), there is an outlier. Exactly calculating the Shapley values is time intensive, in the subset shown there is an outlier that skews the figure and makes it so that viewing the importance distribution excluding the outlier is difficult. For this reason, another shapley summary vas created that looked at different, and larger distribution. This can be seen in Figure (\ref{fig:ShapleySummary2}).

\begin{figure}
    \centering
    \caption[Summary feature importance of a neural network using a larger distribution of data.]{Feature importance of a neural network with model architecture as described in Table \ref{table:gridSearchHyperparamters} and data as described in Table \ref{table:trainDataExample}. Generally multiple factors influence the prediction, with the station elevation being highly influential. Elevation data is excluded when working with Shapley values, as the contribution of each elevation. In contrast to Figure (\ref{fig:ShapleySummary}), the distribution doesn't have as extreme outliers. This means that more details can be seen in the figure. The X-axis shows the influence of feature values on the model. The color gradient shows the value of each feature. As an example, there is a very red value for station elevation (top line, all the way to the left). This means that in this instance, the station elevation contributed around -0.25 to the final output and that the station had an elevation significantly above average.}
    \includegraphics[scale = 0.6]{Figures/shap_plots/summary_plot_190924_.png}
    \label{fig:ShapleySummary2}
\end{figure}

The Figures (\ref{fig:ShapleySummary}) and (\ref{fig:ShapleySummary2}) show the summary for a model trained only on the features shown and not on landscape elevation of the surrounding area. This is done as the number of points there is too high to display in one figure (70 total points). Looking specifically at Figure (\ref{fig:ShapleySummary2}), the station elevation is most influential and the Richardson number's influence is very low. Most of the feature values bunch up in the middle, while the station elevation is elongated compared to other features. Station elevation seems to have two bunches on either side of 0. Overall, the values seem to be skewed to the right of 0. This would be expected as the predicted values are expected to be in the range of 1.2-2, or at least always above 1 by definition.

Finally, for the Shapley values, looking that all the data there are again outliers that skew the data so that spotting general distribution is difficult. This can be seen in Figure (\ref{fig:ShapleySummary3}).

\begin{figure}
    \centering
    \caption[Summary feature importance of a neural network using entire dataset.]{Feature importance of a neural network with model architecture as described in Table \ref{table:gridSearchHyperparamters} and data as described in Table \ref{table:trainDataExample}. The distribution seems to be the same as before, discounting the outliers.}
    \includegraphics[scale = 0.6]{Figures/shap_plots/summary_plot_190924_full_10ms.png}
    \label{fig:ShapleySummary3}
\end{figure}

@@

\begin{figure}
    \centering
    \caption[Summary feature importance of a neural network only looking at AWS at Akrafjall.]{Feature importance of a neural network with model architecture as described in Table \ref{table:gridSearchHyperparamters} and data as described in Table \ref{table:trainDataExample}. This plot only looks at datapoints from Akrafjall. This seems to show the same distribution as previous summary plots. Station elevation is influential and Richardson number has no impact.}
    \includegraphics[scale = 0.6]{Figures/shap_plots/summary_plot_31572.png}
    \label{fig:ShapleySummaryAkrafjall}
\end{figure}

\begin{figure}
    \centering
    \caption[Summary feature importance of a neural network only looking at AWS at Almannaskarð.]{Feature importance of a neural network with model architecture as described in Table \ref{table:gridSearchHyperparamters} and data as described in Table \ref{table:trainDataExample}. This plot only looks at datapoints from Almannaskarð. This seems to show the same distribution as previous summary plots. Station elevation is influential and Richardson number has no impact.}
    \includegraphics[scale = 0.6]{Figures/shap_plots/summary_plot_35553.png}
    \label{fig:ShapleySummaryAlmannaskarð}
\end{figure}

\begin{figure}
    \centering
    \caption[Summary feature importance of a neural network only looking at AWS at Ásgarðsfjall.]{Feature importance of a neural network with model architecture as described in Table \ref{table:gridSearchHyperparamters} and data as described in Table \ref{table:trainDataExample}. This plot only looks at datapoints from Ásgarðsfjall. This seems to show the same distribution as previous summary plots. Station elevation is influential and Richardson number has no impact.}
    \includegraphics[scale = 0.6]{Figures/shap_plots/summary_plot_6745.png}
    \label{fig:ShapleySummaryAsgarðsfjall}
\end{figure}

\begin{figure}
    \centering
    \caption[Summary feature importance of a neural network only looking at AWS at Háahlíð.]{Feature importance of a neural network with model architecture as described in Table \ref{table:gridSearchHyperparamters} and data as described in Table \ref{table:trainDataExample}. This plot only looks at datapoints from Háahlíð. This seems to show the same distribution as previous summary plots. Station elevation is influential and Richardson number has no impact.}
    \includegraphics[scale = 0.6]{Figures/shap_plots/summary_plot_1470.png}
    \label{fig:ShapleySummaryHaahlid}
\end{figure}

\begin{figure}
    \centering
    \caption[Summary feature importance of a neural network only looking at AWS at Keflavíkurflugvöllur.]{Feature importance of a neural network with model architecture as described in Table \ref{table:gridSearchHyperparamters} and data as described in Table \ref{table:trainDataExample}. This plot only looks at datapoints from Keflavíkurflugvöllur. This seems to show the same distribution as previous summary plots. Station elevation is influential and Richardson number has no impact.}
    \includegraphics[scale = 0.6]{Figures/shap_plots/summary_plot_1350.png}
    \label{fig:ShapleySummaryKeflavikurflugvollur}
\end{figure}

@@

>>>>

\begin{table}[h]
    \caption[Model results for different AWSL]{MAPE for each average wind speed limit with and without landscape elevation in a 30° sector around the point of interest into the direction of the reanalysis wind. The influence of adding elevation data seems to reduce the error. The percentage error is higher for lower wind speeds and thus observing the error for different lower bound of wind speed will produce different results. This lower bound is determined using the reanalysis wind speed at 15 meters ($ws_{15}$).}
    \label{table:results}
    \centering
    \begin{tabular}{lccc}
        \toprule
        \textbf{AWSL} & \textbf{MAPE} & &\\ 
        $[m/s]$ & \textit{Baseline} &  \textit{Without DEM} & \textit{With DEM} \\\hline
        $\geq 0$ & 39.2\% & 19.7\% & 18.6\% \\
        $\geq 5$ & 28.1\% & 15.9\% & 14.1\%\\
        $\geq 10$ & 23.9\% & 13.6\% & 11.5\%\\
        $\geq 15$ & 23.2\% & 13.0\% & 10.6\%\\
        $\geq 20$ & 24.7\% & 13.6\% & 11.1\%\\
        $\geq 25$ & 27.7\% & 15.7\% & 12.9\%\\
        \bottomrule
    \end{tabular}
\end{table}

%m/s & $ws_{15}$ & f & $ws_{15}$ & f & $ws_{15}$ & f\\\hline
%$\geq 0$ & 39.2\% & 39.2\% & 19.7\% & 15.9\% & 18.6\% & 15.3\%\\
%$\geq 5$ & 28.1\% & 14.8\% & 15.9\% & 9.9\% & 14.1\% & 9.4\%\\
%$\geq 10$ & 23.9\% & 11.1\%& 13.6\% & 7.6\% & 11.5\% & 7.4\%\\
%$\geq 15$ & 23.2\% & 9.3\% & 13.0\% & 7.0\% & 10.6\% & 6.4\%\\
%$\geq 20$ & 24.7\% & 8.2\% & 13.6\% & 6.3\% & 11.1\% & 5.8\%\\
%$\geq 25$ & 27.7\% & 7.3\% & 15.7\% & 5.8\% & 12.9\% & 5.5\% \\
%7.5 & 25.3\% & 12.6\% & 14.5\% & 8.6\% & 12.5\% & 8.6\%\\
%12.5 & 23.3\% & 10.1\% & 13.1\% & 7.1\% & 11.0\% & 6.9\%\\
%17.5 & 23.7\% & 8.7\% & 12.8\% & 6.4\% & 10.5\% & 6.1\%\\
%22.5 & 26.1\% & 7.8\% & 14.5\% & 6.1\% & 12.6\% & 5.7\%\\

This is some improvement upon the baseline error, with a decrease in error from 23.9\% to 13.6\% and 10.6\% for the baseline, model without DEM and with model with DEM for 10 m/s cutoff. The power generated by wind mill increases with wind speed cubed\cite{wind_power}. The highest wind gusts in Iceland are around 70 m/s. Knowing the gust factor with half as much error as before can allow better anticipation and thus spare turbines for high wind gusts. Another way to look at the error improvement is by station. No location data was directly included in the training data. In Table \ref{table:station_mae_distribution} the mean absolute error of predicted average wind speed and measured average wind speed can be seen for the extreme values. A question to ask is it possible to achieve better results when only looking at a single station?

\begin{table}[h]
    \caption[Model result by station]{The MAPE results for selected stations of interest, both when training for the specific site and when the stations are a part of the general data. For every station the AWSL is set at 10 m/s. In training for a single station at a time, some site specific information can be gauged. This does not mean that the a better result can be reached for that site. Factors such as the number of datapoints at given location can significantly impact the result. This table uses the measured wind speed to determine the cut off for data points. This leads to some data leakage and an increased performance compared to using the reanalysis CARRA speed for cut off.}
    \label{table:specific_sites}
    \centering
    \begin{tabular}{lccc}
        \textbf{Station name} & \textbf{Number of measurement points} & \multicolumn{2}{c}{\textbf{MAPE}}\\
        & & General training & Site training\\\hline
        Akrafjall & 42,791 & 18.6\% & 93.7\%\\
        Almannaskarð & 4,014 & 12.2\% & 86.7\%\\
        Ásgarðsfjall & 15,121 & 9.1\% & 9.4\%\\
        Jökulheimar & 17,176 & 7.7\% & 7.7\%\\
        Sandbúðir & 18,718 & 6.8\% & 6.4\%\\
        Stórholt & 35,126 & 7.1\% & 29.2\% \\
        Þúfuver & 19,538 & 6.4\% & 6.8\%\\
    \end{tabular}
\end{table}

@@

\begin{figure}[h]
    \centering
    \includegraphics[scale = 0.5]{Figures/errorMap.png}
    \caption[MAPE error distribution of stations shown on a map of Iceland.]{The MAPE error of each station in data shown as color gradient circles. That is each station is represented by one circle, with the error value represented as a color gradient from dark blue to yellow. The lowest error at a single station was around 7\% at Garðskagaviti and the highest around 35\% at Seljalandsdalur. It is important to note that the model is trained using a cutoff of 10 m/s and this cutoff point is determined using the reanalysis wind at 15 meters above ground.}
    \label{fig:errorMap}
\end{figure}

@@

Table (\ref{table:closed_intervals}) shows no improvement over the values in the last column in Table (\ref{table:results}). As previously mentioned, the gust factor decreases with increasing wind speed and thus, training on intervals and lessening this effect might be expected to give better results. This does not seem to be the case. Some interesting sites to look closer at for drivers might include places like Kjalarnes, Ingólfsfjall, Þrengsli and others. These can be seen in Table (\ref{table:specific_sites}).

\begin{table}[h]
    \caption[Model result by stations of interest]{The MAPE results of different stations for several stations of interest, both when training for the specific site and when the stations are a part of the general data. For every station the AWSL is set at 10 m/s. In training for a single station at a time, some site specific information can be gauged. This does not mean that the a better result can be reached for that site. Factors such as the number of datapoints at given location can significantly impact the result. This table uses the measured wind speed to determine the cut off for data points. This leads to some data leakage and an increased performance compared to using the reanalysis CARRA speed for cut off.}
    \label{table:specific_sites}
    \centering
    \begin{tabular}{l | cc}
        \toprule
        \textbf{Station name} & \multicolumn{2}{c}{\textbf{MAPE}}\\
         & Baseline & Model\\
         \midrule
        Fáskrúðsfjörður & 28.2\% & 21.8\%\\
        Ingólfsfjall & 30.0\% & 19.6\%\\
        Kjalarnes & 20.7\% & 13.5\% \\
        Sandskeið & 13.0\% & 10.2\%\\
        Seyðisfjörður & 32.1\% & 23.0\%\\
        Þjórsárdalur & 12.2\% & 11.4\%\\
        Þrengsli & 13.6\% & 11.3\%\\
        \bottomrule
    \end{tabular}
\end{table}

\begin{figure}[h]
    \centering
    \includegraphics[scale = 0.6]{Figures/shap_plots/waterfall_plot.png}
    \caption[Feature importance for a single observation of a neural network.]{Feature importance of a neural network with model architecture as described in Table \ref{table:gridSearchHyperparamters} and data as described in Table \ref{table:trainDataExample}. In this specific instance the wind direction ($wd_{15}$) has the highest negative influence and the temperature has the highest positive influence. Elevation data is excluded when working with Shapley values, as the contribution of each elevation point is very low and there are very many of them. To see their influence on the model output see Table \ref{table:results}.}
    \label{fig:ShapleyWaterfall}
\end{figure}

In Figure (\ref{fig:ShapleySummary}) the contribution of each feature, excluding the elevation points, can be seen for the model in general (a significant subset of data is used). Looking at Figure (\ref{fig:ShapleySummary}), there is an outlier. Exactly calculating the Shapley values is time intensive, in the subset shown there is an outlier that skews the figure and makes it so that viewing the importance distribution excluding the outlier is difficult. For this reason, another shapley summary vas created that looked at different, and larger distribution. This can be seen in Figure (\ref{fig:ShapleySummary2}).

\begin{figure}
    \centering
    \includegraphics[scale = 0.6]{Figures/shap_plots/summary_plot.png}
    \caption[Summary feature importance of a neural network.]{Feature importance of a neural network with model architecture as described in Table \ref{table:gridSearchHyperparamters} and data as described in Table \ref{table:trainDataExample}. We can see that generally multiple factors influence the prediction, with the station elevation being highly influential. There is seemingly one outlier for the Richardson number, which usually has very little influence. Elevation data is excluded when working with Shapley values, as the contribution of each elevation point is very low and there are very many of them. To see their influence on the model output see Table \ref{table:results}.}
    \label{fig:ShapleySummary}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[scale = 0.6]{Figures/shap_plots/summary_plot_190924_.png}
    \caption[Summary feature importance of a neural network using a larger distribution of data.]{Feature importance of a neural network with model architecture as described in Table \ref{table:gridSearchHyperparamters} and data as described in Table \ref{table:trainDataExample}. Generally multiple factors influence the prediction, with the station elevation being highly influential. Elevation data is excluded when working with Shapley values, as the contribution of each elevation. In contrast to Figure (\ref{fig:ShapleySummary}), the distribution doesn't have as extreme outliers. This means that more details can be seen in the figure. The X-axis shows the influence of feature values on the model. The color gradient shows the value of each feature. As an example, there is a very red value for station elevation (top line, all the way to the left). This means that in this instance, the station elevation contributed around -0.25 to the final output and that the station had an elevation significantly above average.}
    \label{fig:ShapleySummary2}
\end{figure}

The Figures (\ref{fig:ShapleySummary}) and (\ref{fig:ShapleySummary2}) show the summary for a model trained only on the features shown and not on landscape elevation of the surrounding area. This is done as the number of points there is too high to display in one figure (70 total points). Looking specifically at Figure (\ref{fig:ShapleySummary2}), the station elevation is most influential and the Richardson number's influence is very low. Most of the feature values bunch up in the middle, while the station elevation is elongated compared to other features. Station elevation seems to have two bunches on either side of 0. Overall, the values seem to be skewed to the right of 0. This would be expected as the predicted values are expected to be in the range of 1.2-2, or at least always above 1 by definition. Finally, for the Shapley values, looking that all the data there are again outliers that skew the data so that spotting general distribution is difficult. This can be seen in Figure (\ref{fig:ShapleySummary3}).

\begin{figure}
    \centering
    \includegraphics[scale = 0.6]{Figures/shap_plots/summary_plot_190924_full_10ms.png}
    \caption[Summary feature importance of a neural network using entire dataset.]{Feature importance of a neural network with model architecture as described in Table \ref{table:gridSearchHyperparamters} and data as described in Table \ref{table:trainDataExample}. The distribution seems to be the same as before, discounting the outliers.}
    \label{fig:ShapleySummary3}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[scale = 0.6]{Figures/shap_plots/summary_plot_31572.png}
    \caption[Summary feature importance of a neural network only looking at AWS at Akrafjall.]{Feature importance of a neural network with model architecture as described in Table \ref{table:gridSearchHyperparamters} and data as described in Table \ref{table:trainDataExample}. This plot only looks at datapoints from Akrafjall. This seems to show the same distribution as previous summary plots. Station elevation is influential and Richardson number has no impact.}
    \label{fig:ShapleySummaryAkrafjall}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[scale = 0.6]{Figures/shap_plots/summary_plot_35553.png}
    \caption[Summary feature importance of a neural network only looking at AWS at Almannaskarð.]{Feature importance of a neural network with model architecture as described in Table \ref{table:gridSearchHyperparamters} and data as described in Table \ref{table:trainDataExample}. This plot only looks at datapoints from Almannaskarð. This seems to show the same distribution as previous summary plots. Station elevation is influential and Richardson number has no impact.}
    \label{fig:ShapleySummaryAlmannaskarð}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[scale = 0.6]{Figures/shap_plots/summary_plot_6745.png}
    \caption[Summary feature importance of a neural network only looking at AWS at Ásgarðsfjall.]{Feature importance of a neural network with model architecture as described in Table \ref{table:gridSearchHyperparamters} and data as described in Table \ref{table:trainDataExample}. This plot only looks at datapoints from Ásgarðsfjall. This seems to show the same distribution as previous summary plots. Station elevation is influential and Richardson number has no impact.}
    \label{fig:ShapleySummaryAsgarðsfjall}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[scale = 0.6]{Figures/shap_plots/summary_plot_1470.png}
    \caption[Summary feature importance of a neural network only looking at AWS at Háahlíð.]{Feature importance of a neural network with model architecture as described in Table \ref{table:gridSearchHyperparamters} and data as described in Table \ref{table:trainDataExample}. This plot only looks at datapoints from Háahlíð. This seems to show the same distribution as previous summary plots. Station elevation is influential and Richardson number has no impact.}
    \label{fig:ShapleySummaryHaahlid}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[scale = 0.6]{Figures/shap_plots/summary_plot_1350.png}
    \caption[Summary feature importance of a neural network only looking at AWS at Keflavíkurflugvöllur.]{Feature importance of a neural network with model architecture as described in Table \ref{table:gridSearchHyperparamters} and data as described in Table \ref{table:trainDataExample}. This plot only looks at datapoints from Keflavíkurflugvöllur. This seems to show the same distribution as previous summary plots. Station elevation is influential and Richardson number has no impact.}
    \label{fig:ShapleySummaryKeflavikurflugvollur}
\end{figure}

@@

A simpler way to look at feature importance is to create models that are trained on and use to predict different sets of parameters. The results of such a comparison can be seen in Table \ref{table:setsOfParams}.

\begin{table}[h]
    \caption[Model results for different sets of parameters.]{...}
    \label{table:setsOfParams}
    \centering
    \begin{tabular}{lc}
        \toprule
        \textbf{Model parameters} & \textbf{MAPE}\\ 
        \midrule
        Baseline constant & 23.9\%\\
        $ws_{15}$ & 17.2\%\\
        $[ws_{15}, t_{15}, p_{15}, wd_{15}]$ 16.7\% & ---\\
        $[ws_{15}, t_{15}, p_{15}, wd_{15}, ASL, twd]$ & ---\\
        $[ws_{15}, t_{15}, p_{15}, wd_{15}, ASL, twd, N, Ri]$ & ---\\
        $[ws_{15}, t_{15}, p_{15}, wd_{15}, ASL, twd, N, Ri]$ + DEM & ---\\
        $[ws_{15,250,500}, t_{15,250,500}, p_{15,250,500}, wd_{15,250,500}, ASL, twd, N, Ri]$
        \bottomrule
    \end{tabular}
\end{table}
